---
title: "STAT511 HW1"
author: "Amy Fox"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


# Question 1: Describe the data in Problem 3.3

## Code
```{r}
# read in the data
lumber <- read.csv("../Data/OTT_Final/ASCII-comma/CH03/ex3-30.txt", quote = "'")

# look at the structure of the data
str(lumber)

# print the histogram
hist(lumber$Number, col = "grey", main = "Question 1A: Lumber Histogram")

# find the average of the lumber data
mean(lumber$Number)

# find the median of the lumber data
median(lumber$Number)

```

## Answer
A) The histogram is shown above.
B) The mean for the lumber data is `r mean(lumber$Number)`. The median for the lumber data is `r median(lumber$Number)`. 
C) Based on the histogram, it appears as though the data is normally distributed because the data appears as though there is centered around a number (7 in this plot). Also, the rest of the data does not seem to be left- or right- skewed.

# Question 2: Describe the data in Problem 3.7


## Code
```{r}
# read in the survival data
survival <- read.csv("../Data/OTT_Final/ASCII-comma/CH03/ex3-7.txt", quote = "'")

# look at the structure of the data
str(survival)

# print the median and standard deviation of the Standard therapy
mean(survival$StandardTherapy)
sd(survival$StandardTherapy)

# print the median and standard deviation of the New therapy
mean(survival$NewTherapy)
sd(survival$NewTherapy)

# tidy data for median and mean of both of the therapies
survival_tidy <- survival %>%
  gather(key = Treatment, value = data) %>%
  group_by(Treatment) %>%
  summarise(mean = mean(data),
            sd = sd(data))

# print the boxplot
boxplot(survival, main= "Question 2B: Boxplot of Therapies")

```

## Answer
A) The Standard Therapy mean is `r mean(survival$StandardTherapy)` and standard deviation is `r sd(survival$StandardTherapy)`. The New Therapy mean is `r mean(survival$NewTherapy)` and standard deviation is `r sd(survival$NewTherapy)`.

B) The boxplot is shown above.

