---
title: "Stat511 Hw10"
author: "Amy Fox"
date: "11/19/2019"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
**Bacillus Calmette-Guerin (BCG) is a vaccine for preventing tuberculosis. For this question, we will examine data from 3 studies (Vandiviere et al 1973, TPT Madras 1980, Coetzee & Berjak 1968). The data is summarized below.**

**A. Calculate the odds ratio (corresponding to TB_pos for Trt vs Ctrl) for each study separately. (4 pts)**
```{r}
library(lawstat)
BCG_response <- array(c(619, 2537, 10, 8,
        87892, 87886, 499, 505,
        7232, 7470, 45, 29),
      dim = c(2, 2, 3),
      dimnames = list(Treatment = c("Ctrl", "Trt"),
                      Response = c("TB_neg", "TB_pos"),
                      Study = c("1", "2", "3")))

cmh.test(BCG_response)
```
OR Study 1: 0.19519

OR Study 2: 1.01210

OR Study 3: 0.62391

**B. Use the Breslow-Day test to test for equality of odds ratios across the 3 studies. State your p-value and conclusion. Can we conclude that the odds ratios are equal across the 3 studies? Based on this test, should be combine information across studies? (4 pts)**

**A note about the BCG vaccine from Wikipedia:**

**The most controversial aspect of BCG is the variable efficacy found in different clinical trials that appears to depend on geography. Trials conducted in the UK have consistently shown a protective effect of 60 to 80%, but those conducted elsewhere have shown no protective effect, and efficacy appears to fall the closer one gets to the equator.**

```{r warning = FALSE, message = FALSE}
library(metafor)
cmh_bd_BCG <- rma.mh(ai = BCG_response[1, 1, ], bi =  BCG_response[1, 2, ],
       ci = BCG_response[2, 1, ], di = BCG_response[2, 2, ])

p_value <- cmh_bd_BCG$BDp
```
p-value = `r p_value`

Because the p-value is less than $\alpha$ = 0.05, we reject H0 and assume that the odds ratios are not equal across the 3 studies. Therefore, we should not combine information across studies.

# Question 2

**Problem 10.36 involves bomb hits during WWII. Bomb hits were recorded in n = 576 grids in a map of a region of South London.**

**A. Find the sample mean  ($\hat{\mu}$) bomb hits per grid.**
```{r}
bomb_hits = c(0, 1, 2, 3, 4) # Y aka events
grids = c(229, 211, 93, 35, 8) # obs aka number of events aka units

cbind(bomb_hits, grids)

mu_hat <- sum(bomb_hits*grids)/sum(grids)

```
Sample mean: `r mu_hat`


**B. Use the GOF test to test whether the number of bomb hits per grid follows the Poisson distribution. Calculate the GOF test statistic, df, p-value and give a conclusion using $\alpha$ = 0.05. (6 pts)**
```{r}
# calculate Poisson probability
pois_dist <- dpois(bomb_hits, mu_hat)

# change so that the poisson adds up to 1
sum_pois <- sum(pois_dist)
pois_dist[5] <- 1-sum(pois_dist[1:4])

# expected values
Exp <- pois_dist*576

# contributions of Chi-squared
X2 <- (grids-Exp)^2/Exp

cbind(bomb_hits, grids, pois_dist, Exp, X2)

Chi_Sq_TS <- sum(X2)


df = 5-2
pval <- 1-pchisq(Chi_Sq_TS, df)
pval
```

GOF Test statistic: `r Chi_Sq_TS`

df: `r df`

p-value: `r pval`

Because the p-value > $\alpha$ = 0.05, we fail to reject H0 and conclude that there is no evidence against Poisson aka it follows a poisson distribution.

# Question 3
**The data “PoissonData.csv” gives observations Y (counts or events) for n = 50 (units) generated from the Poisson distribution (using the rpoiss() function).**

**A. Calculate the sample mean and sample standard deviation. Also construct a histogram and qqplot of the data and include them in your assignment. (4 pts)**

**NOTE: Because the data comes from the Poisson distribution, you should find that the mean and the sample variance (s^2^) are close. However, you should also find from the histogram and qqplot that the data looks approximately normal.**
```{r}

library(ggplot2)
poisson_data <- read.csv("../Data/PoissonData.csv")

# mean
poisson_mean <- mean(poisson_data$Y)

# standard deviation
poisson_sd <- sd(poisson_data$Y)

# histogram
ggplot(poisson_data, aes(Y)) +
  geom_histogram(fill = "pink", bins = 7) +
  ggtitle("Poisson Distribution")

# qqplot
qqnorm(poisson_data$Y, main = "QQplot of Poisson Distribution"); qqline(poisson_data$Y)

```

The mean is `r poisson_mean`.

The standard deviations is `r poisson_sd`.

Based on the histogram and qqplot, the data looks pretty normal. 

**B. Give a standard t-based 95% confidence interval for $\mu$.**
```{r}
t.test(poisson_data$Y)
```
CI: (46.29, 50.47)

**C. Following the example on CH10 Slide 106 (Death by Mule Kick CI), construct a 95% confidence interval for $\mu$ based on the normal approximation to the Poisson distribution. (4 pts) In order to do this, you will start by constructing a CI on the total number of events, then divide by the number of units.**

**NOTE: The CIs from parts B and C should be similar.**

$$(y \pm z_{0.025} *  \sqrt{y})/n$$
```{r}
y = sum(poisson_data$Y) 
s = sd(poisson_data$Y)

CI_low <- (y - 1.96*sqrt(y))/50
CI_high <- (y + 1.96*sqrt(y))/50
```

CI: (`r CI_low`, `r CI_high`) 