---
title: "STAT511 Hw2"
author: "Amy Fox"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## Assume that Z has a standard normal distribution. Compute the following.

## A. P(Z $\le$ 0.64)
```{r}
pnorm(0.64)
```
## B. P(Z $\le$ -0.37)
```{r}
pnorm(-0.37)
```

## C. P(Z > 1.24)
```{r}
pnorm(1.24, lower.tail = FALSE)
```

## D. P(-0.37 $\le$ Z $\le$ 1.15)
```{r}
pnorm(1.15) - pnorm(-0.37)
```

## E. Find the value z such that P(Z $\le$ z) = 0.3300

```{r}
qnorm(0.3300)
```

## F. Find the value z such that P(Z > z) = 0.1841

```{r}
qnorm(0.1841, lower.tail = FALSE)
```

# Question 2
## Assume that Y has a normal distribution with mean 5.4 and standard deviation 0.2. Compute the following.

## A. P(Y $\le$ 5.7)
```{r}
pnorm(5.7, mean = 5.4, sd = 0.2)
```

## B. P(Y > 5.3)

```{r}
pnorm(5.3, mean = 5.4, sd = 0.2, lower.tail = FALSE)
```

## C. P(5.2 $\le$ Y $\le$ 5.5)

```{r}
pnorm(5.5, mean = 5.4, sd = 0.2) - pnorm(5.2, mean = 5.4, sd = 0.2)
```

## D. Find the value y such that P(Y $\le$ y) = 0.85.

```{r}
qnorm(0.85, mean = 5.4, sd = 0.2)
```

# Question 3
## Let Y have a skewed distribution with $\mu$ = 80 and $\sigma$ = 5. Suppose a random sample of size n=100 is drawn from the population.

## A. Give an interval with the property that at least 75% of the data will be in that interval. What rule did you use to determine the interval?

Chebyshev's rule is used to determine the interval where for any distribution (non-normal), at least 75% of the data will lie within $\bar{y}$ ±  2s. In this case, the data will lie within 80 ± 10, therefore the internval is (70, 90)

## B. Describe the distribution of $\bar{y}$. Give the mean, standard deviation and shape of the distribution. (3 pts)

```{r}
# standard deviation
5/sqrt(100)
```

The mean is 80. For non-normal data, the standard deviation is sigma/sqrt(n). The standard deviation is `r 5/sqrt(100)`, and the shape of the distribution is approaching normal because according to the Central Limit Theorem, as n increases, the distribution becomes close to normal.


# Question 4
## A random sample of n=25 seeds from a particular bean population is obtained. The weight (g) of each seed is recorded. The data is available from Canvas as “Seeds.csv”.

## Reminders: (1) Use read.csv() to import the data. (2) Use str() to check the data after importing. (3) Use $ or with() to access the Weight column!

## A. Construct a histogram of the data. Also give the sample mean and sample standard deviation. (3 pts)
```{r warning = FALSE}
library(ggplot2)
seed_data <- read.csv("../Data/Seeds.csv")

hist(seed_data$Weight)

# mean
mean(seed_data$Weight)

# standard deviation
sd(seed_data$Weight)
```


## B. Give a 95% confidence interval for $\mu$ (population mean seed weight).
For a 95% confidence interval, $\alpha$ = 0.05 and $\alpha$/2 = 0.025.
First t-$\alpha$/2 must be calculated. df = n-1 where n = 25
```{r}
# this is t-alpha/2
qt(1-0.025, df = 25-1)

mean(seed_data$Weight) + qt(1-0.025, df = 25-1)*(sd(seed_data$Weight)/sqrt(25))

mean(seed_data$Weight) - qt(1-0.025, df = 25-1)*(sd(seed_data$Weight)/sqrt(25))

# or I can do this...
t.test(seed_data$Weight)
```
The CI is (`r mean(seed_data$Weight) - qt(1-0.025, df = 25-1)*(sd(seed_data$Weight)/sqrt(25))`, `r mean(seed_data$Weight) + qt(1-0.025, df = 25-1)*(sd(seed_data$Weight)/sqrt(25))`).

## C. Interpret your confidence interval from part B.

I am 95% confident that the true population mean of the seed weight is between 479g and 573g. 

## D. Do you think the CI is valid? In other words, discuss whether assumptions satisfied.

We must assume that this seed data is a random sample with independent observations. Further, the data must be normally distributed or have a large sample size. 

We know that the data is a random sample as stated in the problem. The data also appears to be normally distributed as shown in the histogram.

# Question 5

## Describe how the following affect the width of the confidence interval (assuming everything else is held constant). Answer should be increase, decrease or stays the same.

## A. Sample size increases.
Decrease

## B. Confidence level increases.
Increase

## C. Standard deviation increases.
Increase
