---
title: "Stats511 Hw5"
author: "Amy Fox"
date: "10/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(dplyr)
library(tidyr)
library(coin)
```

# Question 1

**Refer to Problem 6.42 which deals with lung capacity of rats exposed to ozone. Note: For consistency, please calculate the differences as After – Before where needed.**


**A. Calculate the mean and standard deviation for Before and After (separately).**
```{r}
# read in data
lung_capacity_data <- read.csv("../Data/OTT_Final/ASCII-comma/CH06/ex6-42.txt") %>%
  rename(rat = `X.Rat.`,
         before = `X.Before.`,
         after = `X.After.`) %>%
  mutate(diff = after - before,
         sd_diff = sd(diff)) 

lung_capacity_data %>%
  gather(key = "time", value = "capacity", -rat) %>%
  group_by(time) %>%
  summarise(mean = mean(capacity),
            sd = sd(capacity))


```


**B. Are the differences (After – Before for each rat) normally distributed? Support your answer by including a qqplot of differences in your assignment.**
```{r}

qqnorm(lung_capacity_data$after - lung_capacity_data$before, main = "QQ Plot 
       Before - After Data")
qqline(lung_capacity_data$after - lung_capacity_data$before)

shapiro.test(lung_capacity_data$diff)

```
It appears as if the data is mostly normally distributed based on the QQplots, however, there are a few points that could be outliers.  (I also performed a Shapiro Wilks test to confirm that the data is normally distributed. Because the p-value = 0.2689 > alpha = 0.05 we assume that it is normally distributed.)


**C. Is there sufficient evidence to support the research hypothesis that there is a difference in average lung capacity after ozone exposure? Use the paired t-test with $\alpha$ =0.05. Give the hypotheses, test statistic, p-value and conclusion. (4 pts)**

**Hypothesis**

H0: $\mu$ (before) = $\mu$ (after)

HA: $\mu$ (before) $\neq$  $\mu$ (after)

**Test Statistic**

```{r}
# test statistic and p-value
t.test(lung_capacity_data$after, lung_capacity_data$before, paired = T)

# another way to calculate test statistic
d_line <- mean(lung_capacity_data$before) - mean(lung_capacity_data$after)
s_d <- lung_capacity_data$sd_diff[1]
n_d <- length(lung_capacity_data$diff)

d_line/(s_d/sqrt(n_d))

```

Test statistic: 3.885

p-value: 0.0025

Conclusion: Reject H0 because p-value = 0.0025 < $\alpha$ = 0.05. Therefore, the true difference between the means is not equal to 0.


**D. Rerun the test from the previous question using the Wilcoxon Paired (Signed Rank) test. Give your p-value and conclusion. Use the wilcoxsign_test() function from the coin package with distribution = “exact”.**

```{r}
wilcoxsign_test(lung_capacity_data$after ~ lung_capacity_data$before, paired = T, distribution = "exact")
```

Because the p-value = 0.004 < $\alpha$ = 0.05, we reject H0. Therefore, there is most likely a difference between the true mean lung capacity before and after ozone exposure.


# Question 2

**Refer to problem 7.9 which deals with rebound coefficients of baseballs. The summary statistics are provided here for your convenience: n = 40, $\bar{y}$ = 84.798, s = 2.684. The raw data is also available from the Ott & Longnecker companion site as “exp07-9.txt”. Note that Table 7 (chisquare) does not have information for df = 39, so use the qchisq() R function to calculate table values needed for parts C and D.**


**A. Construct a boxplot of the data and include it in your assignment.**
```{r}
baseball_data <- read.csv("../Data/OTT_Final/ASCII-comma/CH07/exp07-9.txt") %>%
  rename(coefficient = `X.coefficient.`)

boxplot(baseball_data$coefficient, main = "Baseball Boxplot", ylab = "Rebound Coefficient")
``` 


**B. Using $\alpha$ = 0.01, test H0: µ $\geq$ 85 vs HA: µ < 85. Give the one-sided p-value and conclusion.**

```{r}
t.test(baseball_data$coefficient, alternative = "less", mu = 85)
```

Because the p-value = 0.318 > $\alpha$ = 0.01, we fail to reject H0. Thus, the true mean is most likely greater than or equal to 85.


**C. Construct a 99% CI for $\sigma$. Note: provide a standard “two-sided” CI here.**

```{r}
alpha = 0.01
n = 40
s = 2.684

chi_1 <- qchisq(alpha/2, df = 39) 
chi_2 <- qchisq(1-(alpha/2), df = 39) 

# one CI bound
sqrt(((n-1)*s^2)/chi_1)

# second CI bound
sqrt(((n-1)*s^2)/chi_2)

```
CI = (2.07, 3.74)


**D. Using $\alpha$ = 0.01, test H0: $\sigma$ $\leq$ 2 vs HA: $\sigma$ > 2. Give your test statistic, rejection rule and conclusion. (4 pts)**
By hand

**Test Statistic**

```{r}
alpha = 0.01
n = 40
s = 2.684
sigma_0 = 2

(n-1)*s^2/(sigma_0^2)

```

Test statistic = 70.2

**Rejection rule**
```{r}
df = n-1

qchisq(1-alpha, df)
```

Rejection region = 62.4

**Conclusion**
Because Chi^2 = 70.2 > 62.4 = Chi^2 alpha, n-1

Reject H0

Therefore, the true standard deviation is greater than 2.