---
title: "Power_analysis_Prelims"
author: "Amy Fox"
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(stats)
```


# Aim 1.1

**Note: This sample size is required because I'm comparing placenta and amniotic fluid - if I'm looking at fetal lungs etc. I would technically only need 1 pregnant mouse per group, because the average litter size is 6-8**

Two-sample power analysis
H0: mu(aerosol) = mu(non-aerosol)
HA: mu(aerosol) = mu(non-aerosol)

n = ?
delta = 5 (conjectured uA-u0)
sd = 2 (conjectured sd)
sig.level = 0.05
power = 0.9
alternative = two.sided
```{r}
power.t.test(power = 0.9, delta = 5, sd = 2, sig.level = 0.05, alternative = "two.sided", type = "one.sample")
```
# Aim 1.2 
For the correlation coefficients:
(online calculator for Spearman specifically, n = 11.5 --> round up to 12)

Using 4 groups 4 mice per group, 5 timepoints
Plot with the smallest 
```{r}
library(pwr)
pwr.r.test(n = 16, r = 0.8, sig.level = 0.05)
pwr.r.test(r = 0.8, sig.level = 0.05, p = 0.9)
```

Redo the calculations because now using one-sample, two-sided Mann-Whitney
H0: mu = 0
HA: mu $\neq$ 0 



```{r}
library(wmwpow)
wmwpowd(n = 5, m = 0, distn = 5, distm = 0, sides = "two.sided")
```

# Aim 2

Power test for 2.1 (testing presence of microbiota)
```{r}
power.t.test(delta = 10, sd = 4, power = 0.9, alternative = "one.sided")
```


OTU counts: 
```{r}
library(readxl)
sample_micro <- read_xlsx("../Data/data_from_microbiome_paper.xlsx")

lm_fit <- lm(OTUs ~ Sample, data = sample_micro)

# withing-group variance
anova(lm_fit)["Residuals", "Mean Sq"]

anova(lm_fit)["group", "Mean Sq"]


C <- sample_micro %>%
  filter(Sample == "C1" | Sample == "C2" | Sample == "C3" | Sample == "C4" | Sample == "C5") %>%
  mutate(within_C = var(OTUs))

B <- sample_micro %>%
  filter(Sample == "B1" | Sample == "B2" | Sample == "B3" | Sample == "B4" | Sample == "B5") %>%
  mutate(within_B = var(OTUs)) 

M <- sample_micro %>%
  filter(Sample == "M1" | Sample == "M2" | Sample == "M3" | Sample == "M4" | Sample == "M5") %>%
  mutate(within_M = var(OTUs)) 

C
B
M

# within variance
mean(361737.3, 114828.3, 43546.3)



power.anova.test(groups = 4, between.var = 3617370, within.var = 361737, sig.level = 0.05, power = 0.9)
```


```{r}
flow_data %>%
 group_by(Organ, Group) %>%
  nest() %>%
  mutate(aov_result = map(data, ~aov(pop_percent ~ Group, data = .x)),
         tukey_result = map(aov_result, TukeyHSD),
         tidy_tukey = map(tukey_result, tidy)) %>%
  unnest(tidy_tukey, .drop = TRUE)
```


Example Data for 1.2
```{r}

library(ggplot2)
# Plots 1: Deposition vs. Dose

plot1_data <- data.frame(Aerosol_Dose = c(rep("1.4 * 10^4", 5), rep("2.8 * 10^4", 5), rep("7.0 * 10^4",5), rep("1.4 * 10^5", 5)),
           Deposition = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)))

library(ggpubr)
ggplot(plot1_data, aes(x = factor(Aerosol_Dose, levels = c("1.4 * 10^4", "2.8 * 10^4", "7.0 * 10^4", "1.4 * 10^5")), y = Deposition, color = Aerosol_Dose)) +
  geom_point() +
  xlab("PM Aerosol Dose (particles/cm^3)") +
  ylab("PM Deposition (particles)") +
  ggtitle("Example Data: Maternal Plasma Deposition at 24 hours (n = 20)")

plot1_data <- data.frame(Aerosol_Dose = c(rep(1.4 * 10^4, 5), rep(2.8 * 10^4, 5), rep(7.0 * 10^4,5), rep(1.4 * 10^5, 5)),
           Deposition = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)))

cor.test(plot1_data$Aerosol_Dose, plot1_data$Deposition, method = "spearman")

# Plots 2: Concentration vs. Time

plot2_data <- data.frame(Timepoint= c(rep(0, 5), rep(24, 5), rep(48,5), rep(7*24, 5), rep(14*24, 5)),
           Concentration = c(runif(5, 10, 20), runif(5, 20, 50), runif(5, 10, 30), runif(5, 5, 10), runif(5, 0,7)))


ggplot(plot2_data, aes(x = Timepoint, y = Concentration, color = Timepoint)) +
  geom_point() +
  xlab("Timepoint (hrs)") +
  ylab("Concentration of PM (particles/mL)") +
  ggtitle("Example Data: Maternal Plasma Concentration Over Time (n = 25)")
```
Aim 2.2
```{r}
plot2.2_counts <- data.frame(Aerosol_Dose = c(rep("1.4E4", 5), rep("2.8E4", 5), rep("7.0E4",5), rep("1.4E5", 5)),
`Lymphocytes` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`B_cells` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`NK_cells` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`NKT_cells` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`Neutrophils` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`Macrophages` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`Dendritic_cells` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000)),
`Monocytes` = c(runif(5, 2000, 5000), runif(5, 3000,6000), runif(5, 4000, 7000), runif(5, 5000,8000))) %>%
  gather(key = "cell_type", value = "count", -Aerosol_Dose) 

# Bar chart with cell numbers and groups
ggplot(plot2.2_counts, aes(x = cell_type, y = count, fill = cell_type)) +
  geom_bar(stat = "identity") +
  facet_wrap(~factor(Aerosol_Dose, levels = c("1.4E4", "2.8E4", "7.0E4", "1.4E5")), nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Cell Type") +
  ylab("Cell Numbers") +
  ggtitle("Example Data: Pup Lung Cells at 3 weeks")

# Macrophage time series with cell numbers and groups
plot2.2_time <- plot2.2_counts 

x = runif(1, 2000, 3000)
y = runif(1, 3000, 4000)
z = runif(1, 4000, 5000)
a = runif(1, 5000, 6000)

plot2.2_time <- data.frame(Aerosol_Dose = c(rep("1.4E4", 4), rep("2.8E4", 4), rep("7.0E4",4), rep("1.4E5", 4)), 
           Time = c(rep(c(0, 3, 5, 8), 4)),
           Macrophages = c(x, x + runif(1, 1000, 2000), x + runif(1, 2000, 3000), x+runif(1, 3000, 4000),
                           y, y + runif(1, 1000, 2000), y + runif(1, 2000, 3000), y+runif(1, 3000, 4000),
                           z, z + runif(1, 1000, 2000), z + runif(1, 2000, 3000), z+runif(1, 3000, 4000),
                           a, a + runif(1, 1000, 2000), a + runif(1, 2000, 3000), a+runif(1, 3000, 4000)))


ggplot(plot2.2_time, aes(x = Time, y = Macrophages, color = factor(Aerosol_Dose, levels = c("1.4E4", "2.8E4", "7.0E4", "1.4E5")))) +
  geom_point() +
  geom_line() +
  labs(color = "Aerosol Dose") +
  xlab("Time (weeks)") +
  ylab("Cell Numbers") +
  ggtitle("Example Data: Pup Lung Macrophages Over Time") 

# correlating with OTU count

plot2.2_diversity <- plot2.2_time %>%
  mutate(OTU_count = rev(Macrophages))

library(ggpmisc)
ggplot(plot2.2_diversity, aes(x = OTU_count, y = Macrophages)) +
  geom_point() +
  geom_smooth(method = "lm", se= FALSE) +
   stat_poly_eq(formula = y~x, 
                aes(label = paste( ..rr.label..)), 
                parse = TRUE,
                label.y = .25,
                label.x = .25) + 
  xlab("OTU Count") +
  ylab("Macrophage Cell Numbers") +
  ggtitle("Example Data: Pup Lung Macrophages vs. Diversity") 

# pie chart 
ggplot(plot2.2_counts, aes(x = "", y = count, fill = cell_type)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start = 0) 
```

Aim 2.3

```{r}

plot2.3_data_cytokines <- data.frame(Aerosol_Dose = c(rep("1.4E4", 5), rep("2.8E4", 5), rep("7.0E4",5), rep("1.4E5", 5)),
`TNF-alpha` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`IL-17` = c(runif(5, 0, 4), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`IL-1beta` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`IL-1alpha` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`IFN-gamma` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`MCP-1` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`MIP-1alpha` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`MIP-1beta` = c(runif(5, 0, 3), runif(5, 1, 5), runif(5, 2, 10), runif(5, 10, 50)),
`IL-10` = c(runif(5, 20, 40), runif(5, 10, 20), runif(5, 2, 7), runif(5, 0, 5)),
`IL-6` = c(runif(5, 20, 40), runif(5, 10, 20), runif(5, 2, 7), runif(5, 0, 5)),
`IL-4` = c(runif(5, 20, 40), runif(5, 10, 20), runif(5, 2, 7), runif(5, 0, 5))) %>%
  gather(key = "cytokine", value = "amount", -Aerosol_Dose) 

library(tidyverse)
plot2.3_data_cytokines 

library(scales)
ggplot(plot2.3_data_cytokines, aes(x = factor(Aerosol_Dose, levels = c("1.4E4", "2.8E4", "7.0E4", "1.4E5")), y = amount, fill = cytokine)) +
  geom_bar(stat = "identity") +
  facet_wrap(~cytokine) +
  xlab("Aerosol Dose") +
  ylab("Concentration of Cytokine") +
  ggtitle("Example Data: Pup Lung Cytokine Concentrations at 3 weeks") +
  theme(legend.position = "none")

```


Caging
```{r}
Number of particles needed
1.4*10^5 * 10^6 * 2*4

```

Sizing for slide count in mm
```{r}

GA = c(7.5, 8.5, 9.5, 14.5)

# Abdominal circumference
AC = -55.75 + 7.52*GA - 0.166*GA^2

# Placental diameter
PD = -11.96 + 2.09*GA - 0.046*GA^2 - 0.0005*GA^3

# Placental thickness (PT)
PT = 4.1 - 1.14*GA + 0.115*GA^2 - 0.0031*GA^3

GA = c(7.5, 8.5, 9.5, 14.5)

data.frame(GA, AC, PT)

18.4/2/3.14
```

